Clustering time series based on probability distributions across temporal granularities

### 2021-11-16 main.Rmd
re-worked clustering, choosing metrics

Different approaches to determining number of clusters are required for varied applications and research objectives. It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g., between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another. Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation". The Cluster separation index ($sindex$) proposed by @Hennig2014-ah is used as a guide to select the number of clusters in Section \ref{sec:validation}. $sindex$ [@R-fpc] is a compromise between minimum cluster separation and average cluster separation and is less sensitive to a single or few ambiguous points. $sindex$ cannot be optimized over the number of clusters, since increasing the number of clusters reduces it. The number of clusters required is determined by the slope going from steep to shallow (an elbow).

>>>>

Determining the number of clusters is typically a difficult task. Many metrics have been defined for choosing clusters. Most metrics for choosing the optimal number of clusters, are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters. Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy] and the separation index ($sindex$) [@Hennig2014-ah]. These are constructed by balancing within-cluster homogeneity and between-cluster separation.

These can all give contradictory suggestions when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying outlying observations falling between clusters). There is no one best metric, perhaps a reason why so many exist.

In this work, we have chosen to use $sindex$. It is a very simple, but effective metric. This is computed by averaging the minimum 10% of distances between observations between clusters, making it relatively robust to nuisance observations. The value of $sindex$ always decreases, and where values drop substantially when an extra cluster are candidates for optimal clustering. The number of clusters corresponding to the value before the drop is the recommendation.
